Ages of Stars and Planets in the Kepler Field Over the First Four Billion Years

### 2024-09-03 papers/paper/ms.tex
update ms

\added{{\it Weakened magnetic braking:} 
If we lived in a
universe in which stellar rotation rates and spot-induced photometric
amplitudes remained {\it constant} after $\approx$3\,Gyr, then older
stars could contaminate the peaks in
Figure~\ref{fig:hist_tgyro}.  This
scenario deserves consideration, given asteroseismic evidence for slowed spin-down rates after
$\approx$4\,Gyr for Sun-like stars
\citep[][]{vanSaders_2016,Hall2021,2024ApJ...962..138S}, and the ``long-period edge''
	in the $P_{\rm rot}$-$T_{\rm eff}$ distribution which
	may pile up over similar timescales \citep{David_2022}.  However, 
	the stalling is thought to happen at $>$5\,Gyr for
	late-G dwarfs and early K dwarfs \citep[e.g.~Figure~4
	by][]{2024ApJ...962..138S}.  This means that it cannot explain the
	2.3\,Gyr peak in the 4400-5400\,K age distribution shown in Figure~\ref{fig:hist_tgyro}. 
	Independently,
	the study by \citet{2022ApJ...937...94M} fitted
	isochrones to a high-quality subset of Kepler data, and found that the
	\citet{McQuillan_2014} sample, discussed in Appendix~\ref{app:mcqonly}, does not
	include many stars older than $\approx$\,4\,Gyr.
These arguments suggest that the peak
in the age distribution is not caused by a pile-up of
much older stars with lethargic spin-down rates, but that it could be
	contaminated by them beyond $\approx$3\,Gyr, especially for stars hotter than the Sun.
}

>>>>

\added{{\it Weakened magnetic braking:} 
If stellar rotation rates and spot-induced photometric amplitudes
remained {\it constant} after $\approx$3\,Gyr, then older stars could
contaminate the peaks in Figure~\ref{fig:hist_tgyro}.  This
possibility warrants consideration, given asteroseismic evidence
suggesting slowed spin-down rates after $\approx$4\,Gyr for Sun-like
stars \citep[][]{vanSaders_2016,Hall2021,2024ApJ...962..138S}, and the
``long-period edge'' in the $P_{\rm rot}$-$T_{\rm eff}$ distribution
which may pile up over similar timescales \citep{David_2022}.
However, this stalling is thought to happen at $>$5\,Gyr for late-G
dwarfs and early K dwarfs \citep[e.g.~Figure~4
by][]{2024ApJ...962..138S}, which means that it cannot account for the
2.3\,Gyr peak in the 4400-5400\,K age distribution shown in
Figure~\ref{fig:hist_tgyro}.}

\added{Additionally, \citet{2022ApJ...937...94M} found that the
\citet{McQuillan_2014} sample, discussed in Appendix~\ref{app:mcqonly},
includes few stars older than $\approx$4\,Gyr, based on
isochrone-fitting to a high-quality subset of Kepler planet hosts.
Some solar twins have also been shown to have rotation periods of
30-50 days
\citep{LorenzoOliveira2019,LorenzoOliveira2020,doNascimento2020},
indicating that a completely stalled spin-down may be unlikely. These
arguments suggest that while the peak in the age distribution is not
primarily caused by an accumulation of much older stars with slowed
spin-down rates, contamination by such stars beyond $\approx$3\,Gyr
remains possible, particularly for stars hotter than the Sun.}
